Powdery mildew yield loss mitigation strategies in mungbean, a meta-analysis.

### 2019-11-04 03_Preliminary_analysis.Rmd
Improved formatting flow and correction of broken code

```{r PA_Libraries, include=FALSE}
library(dplyr)
library(openxlsx)
library(kableExtra)
library(ggplot2)
library(devtools)
library(lme4)
library(theme.usq)
source("R/import_data.R")
#source("./getPPD2.R")
```

@@

@@

   2) Fungicide dose

   3) Row spacing

   4) Host cultivars (probably is a covariate with season due to changing cultivars over time)

   5) Experiment location

   6) Final disease rating

1. Type of Fungicide sprayed
2. In-crop rainfall
3. mungbean cultivar
4. Fungicide dose (we don't expect this to influence as it was only changes slightly)
5. Mean daily temperature in the weeks following spray application
6. Row spacing
7. Two pathogen species cause PM (new discovery)
8. irrigated vs non-irrigated trials
9. Number of fungicide sprays (edited)

### Fungicides

The data plotted below are from the past field trials (2011 - 2017) described in the aforementioned table.
All data are a list of mean values from each treatment applied in each experiment.

When investigating previous studies into powdery mildew fungicide efficacy on mungbean `r length(unique(PM_MB_means$trial_ref))` trials were found.
However, only eleven tebuconazole were used for the following meta-analysis, followed by the next most common fungicide, propiconazole. Sulphur was disregarded due to its inconsistency in efficacy.

```{r Fungicides}
PM_MB_means %>%
   group_by(fungicide_ai, trial_ref) %>%
   summarise() %>%
   count(sort = TRUE) %>%
   rename(Trials = n) %>%
   ggplot(aes(x = reorder(fungicide_ai, Trials), y = Trials)) +
   xlab("Fungicide active ingredient") +
   ylab("N Trials") +
   geom_col() +
   scale_fill_usq() +
   ggtitle(label = "Number of trials in which the\nspecified fungicide was used") +
   scale_colour_usq() +
   theme_usq() +
   coord_flip()
```

Tebuconazole and propiconazole might be able to be pooled as a triazole fungicide treatment. 
Amistar Xtra is a combination of azoxystrobin and cyproconazole.
Custodia is a combination of tebuconazole and azoxystrobin.
Amistar Xtra and Custodia might be able to be pooled given they both contain strobilurin and triazole, however, they contain differing dose ratios (inverted).

Perhaps best way forward is to do an analysis of only the triazoles. Then do another including azoxystrobin as a comparison.

Preliminary analysis of fungicides show there is a difference the effect of fungicide in proportion of yield saved.

@@

Were there any statistical difference between all row spacing or only some?
Can we pool certain row spacing that have no significant difference?
From the preliminary analysis, the graphs seem to imply that if there is a lower overall yield there is no effect of row spacing on yield.
However, if the average yield is more than approximately 0.6 - 1 t/ha then smaller row spacing has the potential to provide greater yield.

# Which row spacing leads to the higher disease severity
PM_MB_means %>%
   filter(year == 2017 |
             year == 2018) %>%
   filter(fungicide == "control") %>%
   ggplot(aes(y = PM_final_severity, x = row_spacing, colour = location)) +
   geom_point() +
   #  geom_smooth()+
   scale_y_continuous(limits = c(-5, 15)) +
   theme_usq()

# Which row spacing leads to the higher yeild potential
PM_MB_means %>%
   filter(year == 2017|
             year == 2018) %>%
   #   filter(Fungicide == "control") %>%
   ggplot(aes(
      y = as.numeric(grain_yield.t.ha.),
      x = row_spacing,
      colour = location
   )) +
   geom_smooth(method = "gam") +
   geom_jitter(width = 0.01) +
   #   scale_y_continuous(limits = c(-5,15))+
   theme_usq()

### Host genotypes

Host genotype may need to be analysed to determine the effect.

@@

Host genotype seems to contribute to the variation in yield. Berken shows higher yields compared to crystal and jade.
```{r yield_vol}
source("R/yield_volatility.r") #function to investigate the volatility in yields

yield_volatility(genotype_by_trial = FALSE, control_only = FALSE)
yield_volatility(genotype_by_trial = FALSE, control_only = TRUE)
yield_volatility(genotype_by_trial = FALSE, control_only = FALSE, location = "Hermitage")

yield_volatility("Crystal",control_only = TRUE)
yield_volatility("Crystal",control_only = FALSE)

yield_volatility("Jade",control_only = TRUE)
yield_volatility("Jade",control_only = FALSE, location = "Hermitage")

@@

#### Best start day for in-crop rainfall

Loop over the dates at the start of the season (to 90 days after planting) to find which date fits the best linear model for grain weight and in-crop rainfall. The code for this process was moved to be contained within it's own script and can be run as a job as it can take a long time. The output of the script is saved with the prefix `lmInSeasonRainfall` followed by the time windows the model is run. For example `lmInSeasonRainfall_20.40_50.80.csv` holds the model results for start days between 20 and 40 days after planting to end window dates between 50 and 80 days after planting.

@@

@@

The approximate period when in-crop rainfall impacts the most on the mean harvest is between 11 - 88 days after planting.

@@

```{r Fungicide_x_nSprays, message=FALSE}
PM_MB_means %>%
   group_by(fungicide, total_fungicide) %>%
   summarise(
      n = length(prop_YG),
      lower_2.5 = quantile(prop_YG, na.rm = T, c(0.025)),
      median = median(prop_YG, na.rm = T),
      mean = mean(prop_YG, na.rm = T),
      upper_97.5 = quantile(prop_YG, na.rm = T, c(0.975))
   ) %>%
   filter(n >= 5) %>% # remove any fungicide groups with less than 5 observations
   arrange(desc(median)) %>%
   kable(caption = "Fungicide effect on proportion of yield gain") %>%
   kable_styling(fixed_thead = T, full_width = T)

PM_MB_means %>%
   filter(total_fungicide == 1 | total_fungicide == 2) %>%
   filter(
      fungicide != "pyrazophos" &
         fungicide != "control" &
         fungicide != "benomyl" &
         fungicide != "Acibenzolar-S-methyl" &
         fungicide != "200g/L azoxystrobin + 80 g/L cyproconazole + adj"
   ) %>%
   ggplot(aes(y = prop_YG, x = fungicide)) +
   facet_grid(rows = vars(total_fungicide)) +
   geom_boxplot(aes(fill = fungicide, alpha = 0.5)) +
   geom_hline(aes(yintercept = 0), size = 0.5) +
   #scale_fill_discrete(labels = c("15-7 days before","same day","one day after","8-9 days after","13-19 days after"))+
   labs(y = "Proportion of grain yield saved",
        title = "Grain yield proportion saved\ngrouped by Fungicide type") +
   theme_usq() +
   theme(legend.position = "none") +
   # coord_cartesian() +
   # scale_x_discrete(
   #    limits = c(
   #       "200 g/l  Tebuconazole + 120 g/l  Azoxystrobin",
   #       "Propiconazole",
   #       "Tebeconazole",
   #       "200g/L Azoxystrobin + 80 g/L Cyproconazole",
   #       "Sulphur",
   #       "Carbendazim",
   #       "Pyraclostrobin",
   #       "Azoxystrobin"
   #    )
   # ) +
   #  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3))+
   coord_flip()
#scale_x_continuous(limits = c(-0.3,0.6))

@@

# Making a table of showing the upper and lower percentiles (95%)
quant_PMy <- PM_MB_means %>%
   filter(Fungicide != "control") %>%
   group_by(total_Fungicide) %>%
   summarise(
      n = length(yield_gain),
      lower_2.5 = quantile(yield_gain, na.rm = T, c(0.025)),
      median = median(yield_gain, na.rm = T),
      mean = mean(yield_gain, na.rm = T),
      upper_97.5 = quantile(yield_gain, na.rm = T, c(0.975))
   )

# Density plot of Number of fungicide applications and the resulting yeild saved (as a proportion)
ggplot(subset(PM_MB_means, Fungicide != "control"), aes(yield_gain)) +
   geom_density(aes(
      group = factor(total_Fungicide),
      fill = factor(total_Fungicide)
   ), alpha = 0.6) +
   geom_vline(xintercept = 0, size = 1) +
   labs(x = "Grain yield saved (t/ha)",
        title = "Density plot of grain yield saved from differing numbers of fungicide applications") +
   theme(
      panel.grid.major = element_line(colour = "grey80"),
      panel.grid.minor = element_line(colour = "grey50"),
      panel.background = element_rect(fill = NA)
   ) +
   geom_vline(
      data = quant_PMy,
      aes(xintercept = lower_2.5,
          colour = factor(total_Fungicide)),
      size = 0.9,
      linetype =  "dashed"
   ) +
   geom_vline(
      data = quant_PMy,
      aes(xintercept = upper_97.5,
          colour = factor(total_Fungicide)),
      size = 0.9,
      linetype =  "dashed"
   ) +
   guides(
      colour = guide_legend(title = "Upper (97.5%) and \nLower (2.5%) percentiles"),
      fill = guide_legend(title = "Number of \nFungicide applications")
   ) +
   scale_x_continuous(limits = c(-0.3, 0.8))

>>>>

```{r PA_Libraries, include=FALSE}
library(tidyverse)
library(openxlsx)
library(kableExtra)
library(ggplot2)
library(devtools)
library(lme4)
library(theme.usq)
source("R/import_data.R")
#source("./getPPD2.R")
```

@@

#remove the variation in different sulphur formulations
unique(PM_MB_means[
   grep("sulphur",PM_MB_means$fungicide),
   "fungicide"])
PM_MB_means[grep("sulphur",PM_MB_means$fungicide),"fungicide"] <- "sulphur"

@@

   2) Fungicide dose (varies only slightly within some fungicide types)

   3) Number of fungicide sprays

   4) Timing of fungicide spray/s relative to first sign of disease (ie. when pathogen is present)

   5) Host cultivars (probably is a covariate with season due to changing cultivars over time)

   6) Experiment location

   7) Row spacing

   8) In-crop rainfall

   9) irrigated vs non-irrigated trials

   10) Mean daily temperature in the weeks following spray application

   11) Final disease rating

   12) Two pathogen species cause PM (new discovery)

### Fungicides
#### Fungicide type
The data plotted below are from `r length(unique(PM_MB_means$trial_ref))` field trials between (`r first(levels(PM_MB_means$year))` - `r last(levels(PM_MB_means$year))`) of which the details are described in [Studies considered in meta-analysis section](#studies-considered-in-meta-analysis).  
All data explored below are a list of mean values from each treatment applied in each experiment. Some experiments have in-season observations of disease incidence and severity, however here we have summarised the data to show only the end of season observations.    
```{r Fungicides}
PM_MB_means %>%
   group_by(fungicide_ai, trial_ref) %>%
   summarise() %>%
   count(sort = TRUE) %>%
   rename(Trials = n) %>%
   ggplot(aes(x = reorder(fungicide_ai, Trials), y = Trials)) +
   xlab("Fungicide active ingredient") +
   ylab("N Trials") +
   geom_col() +
   scale_fill_usq() +
   ggtitle(label = "Number of trials in which the\nspecified fungicide was used") +
   scale_colour_usq() +
   theme_usq() +
   coord_flip()
```

   - Tebuconazole and propiconazole might be able to be pooled as a triazole fungicide treatment.

   - Amistar Xtra and Custodia are both contain strobilurin and triazole, however, they because they contain differing dose ratios (inverted) pooling may not be appropriate.

   - Perhaps best way forward is to do an analysis of only the triazoles/demethylation inhibitors. Then do another including azoxystrobin as a comparison.

The primary focus of this research is the effect of fungicide on mitigating yield loss to disease. To reduce the affect of location and seasonal influences we will investigate the proportion of grain yield saved, relative to the no spray controls.

First lets investigate the difference in the effect of fungicide on proportion of yield saved?

```{r Fungicide_x_nSprays, message=FALSE}
PM_MB_means %>%
   group_by(fungicide, total_fungicide) %>%
   summarise(
      n = length(prop_YG),
      lower_2.5 = quantile(prop_YG, na.rm = T, c(0.025)),
      median = median(prop_YG, na.rm = T),
      mean = mean(prop_YG, na.rm = T),
      upper_97.5 = quantile(prop_YG, na.rm = T, c(0.975))
   ) %>%
   filter(n >= 5) %>% # remove any fungicide groups with less than 5 observations
   arrange(desc(median)) %>%
   kable(caption = "Fungicide effect on proportion of yield gain.") %>%
   kable_styling(fixed_thead = T, full_width = T) %>%
   footnote(general = "Fungicides with less than five observations were omitted from this table")

class(PM_MB_means$fungicide)

PM_MB_means %>%
   mutate(fungicide = as.factor(PM_MB_means$fungicide)) %>%
   filter(total_fungicide == 1 | total_fungicide == 2) %>%
   filter(
      fungicide != "pyrazophos" &
         fungicide != "control" &
         fungicide != "benomyl" &
         fungicide != "Acibenzolar-S-methyl"
   ) %>%
   mutate(fungicide = factor(fungicide, 
                             levels = c("carbendazim", "pyraclostrobin",
                                        "sulphur", "200 g/L azoxystrobin + 80 g/L cyproconazole",
                                        "tebuconazole", "propiconazole", 
                                        "200 g/L tebuconazole + 120 g/L azoxystrobin"))) %>%
   ggplot(aes(y = prop_YG, x = fungicide)) +
   facet_grid(rows = vars(total_fungicide)) +
   geom_boxplot(aes(fill = fungicide, alpha = 0.5)) +
   geom_hline(aes(yintercept = 0), size = 0.5) +
   #scale_fill_discrete(labels = c("15-7 days before","same day","one day after","8-9 days after","13-19 days after"))+
   labs(y = "Proportion of grain yield saved",
        title = "Grain yield proportion saved\ngrouped by Fungicide type") +
   theme_usq() +
   theme(legend.position = "none") +
   # coord_cartesian() +
   # scale_x_discrete(
   #    limits = c(
   #       "200 g/l  Tebuconazole + 120 g/l  Azoxystrobin",
   #       "Propiconazole",
   #       "Tebeconazole",
   #       "200g/L Azoxystrobin + 80 g/L Cyproconazole",
   #       "Sulphur",
   #       "Carbendazim",
   #       "Pyraclostrobin",
   #       "Azoxystrobin"
   #    )
   # ) +
   #  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3))+
   coord_flip()
#scale_x_continuous(limits = c(-0.3,0.6))

@@

#### Number of fungicide sprays

#### Timing of fungicide sprays

### Mungbean cultivars  
Host genotype may need to be analysed to determine the effect.

@@

Host genotype seems to contribute to the variation in yield. Berken shows higher yields compared to crystal and jade.
```{r yield_vol}
# this could be converted into a shiny app as a chunk. Make each input for the function a shiny option
# Outputs the table and graph
source("R/yield_volatility.r") #function to investigate the volatility in yields
# Volitility is the range between the upper and lower 97.5 and 2.5 % quartiles

YV1 <- yield_volatility(genotype_by_trial = FALSE, control_only = FALSE)
YV1[[2]]
YV1[[3]]
YV2 <- yield_volatility(genotype_by_trial = FALSE, control_only = TRUE)
YV2[[2]]

YV1_herm <- yield_volatility(genotype_by_trial = FALSE, control_only = FALSE, location = "Hermitage")
YV1_herm[[2]]

YV1_crystal <- yield_volatility("Crystal",control_only = FALSE)
YV1_crystal[[2]]

YV2_crystal <- yield_volatility("Crystal",control_only = TRUE)
YV2_crystal[[2]]

YV1_herm_jade <- yield_volatility("Jade",control_only = FALSE, location = "Hermitage")
YV2_jade <- yield_volatility("Jade",control_only = TRUE)

Were there any statistical difference between all row spacing or only some?
Can we pool certain row spacing that have no significant difference?  
From the preliminary analysis, the graphs seem to imply if there is a lower overall yield there is no effect of row spacing on yield.
However, if the average yield is more than approximately 0.6 - 1 t/ha then smaller row spacing has the potential to provide greater yield per hectare.

# Which row spacing leads to the higher disease severity
PM_MB_means %>%
   filter(year == 2017 |
             year == 2018) %>%
   filter(fungicide == "control") %>%
   ggplot(aes(y = PM_final_severity, x = factor(row_spacing))) +
   geom_boxplot() +
   #  geom_smooth()+
   scale_y_continuous(limits = c(-5, 15)) +
   theme_usq()+
   ggtitle("Powdery mildew variation between different row spacing")

# Which row spacing leads to the higher yeild potential
PM_MB_means %>%
   filter(year == 2017|
             year == 2018) %>%
   #   filter(Fungicide == "control") %>%
   ggplot(aes(
      y = as.numeric(grain_yield.t.ha.),
      x = row_spacing,
      colour = location
   )) +
   geom_smooth(method = "gam") +
   geom_jitter(width = 0.01) +
   #   scale_y_continuous(limits = c(-5,15))+
   theme_usq()+
   ggtitle("Grain yield results for different row spacing at three locations")

@@

Loop over the dates at the start of the season (to 90 days after planting) to find which date fits the best linear model for grain weight and in-crop rainfall. The code for this process was moved to be contained within it's own script and can be run as a job as it can take a long time. To speed up the computation I split the script into four, ran the jobs and bound the data into one frame. The output of the script is saved with the prefix `lmInSeasonRainfall` followed by the time windows the model is run. For example `lmInSeasonRainfall_20.40_50.80.csv` holds the model results for start days between 20 and 40 days after planting to end window dates between 50 and 80 days after planting.

@@

source("R/AUS_rainfall_function.r")
source("R/rainfall_sum_dat.R")

for(i in rain_dat_sum$trial_ref){
dat_rain <- rain_dat_sum[rain_dat_sum$trial_ref == i,]

   rain_dat_sum[which(rain_dat_sum$trial_ref == i), "rainfall_sum"] <- 
      AUS_rainfall(StartDate = as.Date(dat_rain$planting_date) +
                   lm_rain[which(lm_rain$lm_adj_rsquared == max(lm_rain$lm_adj_rsquared)), "start_day"],
                EndDate = as.Date(dat_rain$planting_date) +
                   lm_rain[which(lm_rain$lm_adj_rsquared == max(lm_rain$lm_adj_rsquared)), "end_day"],
                latlong = as.vector(unlist(dat_rain[, 5:6])))
}

summary(lm(rainfall_sum ~ grain_yield, 
                             data = rain_dat_sum[rain_dat_sum$location != "Hermitage",]))

rain_dat_sum %>%
   ggplot(aes(x = grain_yield, y = rainfall_sum, colour = location))+
   geom_point() + 
   geom_abline(intercept = -167.98, slope = 284.44, colour = "red")

@@

The approximate period when in-crop rainfall impacts the most on the mean harvest is between 14 - 67 days after planting.

@@

@@

# Making a table of showing the upper and lower percentiles (95%)
quant_PMy <- PM_MB_means %>%
   filter(fungicide != "control") %>%
   group_by(total_fungicide) %>%
   summarise(
      n = length(yield_gain),
      lower_2.5 = quantile(yield_gain, na.rm = T, c(0.025)),
      median = median(yield_gain, na.rm = T),
      mean = mean(yield_gain, na.rm = T),
      upper_97.5 = quantile(yield_gain, na.rm = T, c(0.975))
   )

# Density plot of Number of fungicide applications and the resulting yeild saved (as a proportion)
ggplot(subset(PM_MB_means, fungicide != "control"), aes(yield_gain)) +
   geom_density(aes(
      group = factor(total_fungicide),
      fill = factor(total_fungicide)
   ), alpha = 0.6) +
   geom_vline(xintercept = 0, size = 1) +
   labs(x = "Grain yield saved (t/ha)",
        title = "Density plot of grain yield saved from differing numbers of fungicide applications") +
   theme(
      panel.grid.major = element_line(colour = "grey80"),
      panel.grid.minor = element_line(colour = "grey50"),
      panel.background = element_rect(fill = NA)
   ) +
   geom_vline(
      data = quant_PMy,
      aes(xintercept = lower_2.5,
          colour = factor(total_fungicide)),
      size = 0.9,
      linetype =  "dashed"
   ) +
   geom_vline(
      data = quant_PMy,
      aes(xintercept = upper_97.5,
          colour = factor(total_fungicide)),
      size = 0.9,
      linetype =  "dashed"
   ) +
   guides(
      colour = guide_legend(title = "Upper (97.5%) and \nLower (2.5%) percentiles"),
      fill = guide_legend(title = "Number of \nFungicide applications")
   ) +
   scale_x_continuous(limits = c(-0.3, 0.8))
